---
title: "STAT 345 Midterm Project"
font: 12pt
date: "Due March 31"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

> "Our offense is like the Pythagorean Theorem. There is no answer!", _Shaquille O'Neal_

```{r, message=FALSE}
#import necessary packages
library(devtools)
library(ggplot2)
library(tidyverse)
library(maps)
library(mapdata)
library(ggthemes)
library(dplyr)

#Import the nbastatR dataset from github:
#Sys.setenv(GITHUB_PAT = "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")
devtools::install_github("abresler/nbastatR")
library(nbastatR)
#usethis::create_github_token()

#this makes sure buffer is big enough
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)

#tibbles of each year's shots (a tibble is basically a dataframe- a collection of data)
shots97 <- teams_shots(teams = "Milwaukee Bucks", seasons = 1997) 
shots98 <- teams_shots(teams = "Milwaukee Bucks", seasons = 1998)
shots99 <- teams_shots(teams = "Milwaukee Bucks", seasons = 1999)
shots00 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2000)
shots01 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2001)
shots02 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2002)
shots03 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2003)
shots04 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2004)
shots05 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2005)
shots06 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2006)
shots07 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2007)
shots08 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2008)
shots09 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2009)
shots10 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2010)
shots11 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2011)
shots12 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2012)
shots13 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2013)
shots14 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2014)
shots15 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2015)
shots16 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2016)
shots17 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2017)
shots18 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2018)
shots19 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2019)
shots20 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2020)
shots21 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2021)
shots22 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2022)
shots23 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2023)

#combine each year's shot data into one tibble
MKE.shots<- bind_rows(shots97, shots98, shots99, shots00, shots01, shots02, shots03, shots04, shots05, shots06, shots07, shots08, shots09, shots10, shots11, shots12, shots13, shots14, shots15, shots16, shots17, shots18, shots19, shots20, shots21, shots22, shots23)


###########court design#############

#creating a court by first using the dataset created above
shot.chart <- MKE.shots %>% 
  #divide location coordinates by 10, color by make/miss
  ggplot(mapping=aes(locationX/10, locationY/10, color=isShotMade)) + 
  #makes points more transparent
  geom_point(alpha=.4) + 
  
  #Lines of court
  geom_segment(aes(22,-4,xend=22, yend=10)) + 
  geom_segment(aes(-25, -4, xend=25, yend=-4)) +
  geom_segment(aes(-22, -4, xend= -22, yend=10))+
  geom_segment(aes(-6,-4, xend=-6, yend=19)) + 
  geom_segment(aes(6, -4, xend=6, yend=19)) + 
  geom_segment(aes(-25,-4,xend=-25, yend=43)) + 
  geom_segment(aes(25,-4, xend=25, yend=43)) + 
  geom_segment(aes(-8, 19, xend=8, yend=19)) + 
  geom_segment(aes(-8, -4, xend=-8, yend=19)) + 
  geom_segment(aes(8,-4, xend=8, yend=19)) + 
  #free throw to 3 pt arch
  geom_curve(aes(-8, 19, xend=8, yend=19), curvature=-0.35) + 
  #3 pt arch
  geom_curve(aes(-22, 10, xend=22, yend=10), curvature= -0.3)+
  #renames legend title, labels, and colors
  scale_color_discrete(name="Is Shot Made?", labels = c("No","Yes"), type=c("dark gray", "dark green")) +
  #delete x label
  xlab(NULL) +
  #delete y label
  ylab(NULL) + 
  #set title of graph
  ggtitle("Milwaukee Bucks Shots") + 
  #display a graph of shot data for each year
  facet_wrap(~yearSeason, nrow=5)

#prints shot chart
shot.chart

```
   
   
   
   
```{r}

giannis18 <- teams_shots(teams= "Milwaukee Bucks", seasons=2018) %>% 
  filter(., namePlayer=="Giannis Antetokounmpo")
giannis19 <- teams_shots(teams= "Milwaukee Bucks", seasons=2019) %>% 
  filter(., namePlayer=="Giannis Antetokounmpo")
giannis20 <- teams_shots(teams= "Milwaukee Bucks", seasons=2020) %>% 
  filter(., namePlayer=="Giannis Antetokounmpo")
giannis21 <- teams_shots(teams= "Milwaukee Bucks", seasons=2021) %>% 
  filter(., namePlayer=="Giannis Antetokounmpo")
giannis22 <- teams_shots(teams= "Milwaukee Bucks", seasons=2022) %>% 
  filter(., namePlayer=="Giannis Antetokounmpo")
giannis23 <- teams_shots(teams= "Milwaukee Bucks", seasons=2023) %>% 
  filter(., namePlayer=="Giannis Antetokounmpo")

#combine each year's shot data into one tibble
giannis.shots<- bind_rows(giannis18, giannis19, giannis20, giannis21, giannis22, giannis23)

giannis.shots



giannis.chart <- giannis.shots %>% 
  #divide location coordinates by 10, color by make/miss, shape by make/miss
  ggplot(mapping=aes(locationX/10, locationY/10, color=isShotMade)) + 
  #makes points more transparent
  geom_point(alpha=.4) + 
  
  #Lines of court
  geom_segment(aes(22,-4,xend=22, yend=10)) + 
  geom_segment(aes(-25, -4, xend=25, yend=-4)) +
  geom_segment(aes(-22, -4, xend= -22, yend=10))+
  geom_segment(aes(-6,-4, xend=-6, yend=19)) + 
  geom_segment(aes(6, -4, xend=6, yend=19)) + 
  geom_segment(aes(-25,-4,xend=-25, yend=43)) + 
  geom_segment(aes(25,-4, xend=25, yend=43)) + 
  geom_segment(aes(-8, 19, xend=8, yend=19)) + 
  geom_segment(aes(-8, -4, xend=-8, yend=19)) + 
  geom_segment(aes(8,-4, xend=8, yend=19)) + 
  #free throw to 3 pt arch
  geom_curve(aes(-8, 19, xend=8, yend=19), curvature=-0.35) + 
  #3 pt arch
  geom_curve(aes(-22, 10, xend=22, yend=10), curvature= -0.3)+
  #renames legend title, entries, and colors
  
  scale_color_discrete(name="Is Shot Made?", labels = c("No","Yes"), type=c("dark gray", "dark green")) +
  #delete x label
  xlab(NULL) +
  #delete y label
  ylab(NULL) + 
  #set title of graph
  ggtitle("Giannis Antetokounmpo Shots") + 
  #display shot data by year
  facet_wrap(~yearSeason, nrow=5)

giannis.chart


```
   
Analyzing Giannis Antetokounmpo's shots from 2018-2023, it appears that his game shifted from shooting a variety of mid-range jumpers to focusing heavily on driving to the hoop to shoot layups or dunk. This trend of Giannis's appears to be the driving reason behind the Bucks's overall increase in made layups in 2023. Another trend in Giannis's shots is that he makes more shots on the right-hand side of the court. This is because he is right-handed, so it is natural that his right hand would be his better ball-handling hand, especially for a forward that doesn't have the strongest ball-handling abilities. Thus, the Bucks can expect that defense will try to force Giannis to go to the left side of the court, since that's where he is historically weakest.